Transactional Memory Concurrency Verification with Landslide

### 2018-03-15 htm/paper.tex
pleease look at this picture. see? you colored it! q d e

\begin{figure}[t]
	\begin{center}
		\begin{tabular}{ll}
		%\texttt{void count() \{} \\
		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
		1 & \texttt{if ((status = \_xbegin()) == SUCCESS) \{} \\
		2 & \texttt{~~~~x++;} \\
		3 & \texttt{~~~~\_xend();} \\
		4 & \texttt{\} else \{} \\
		5 & \texttt{~~~~mutex\_lock(\&m);} \\
		6 & \texttt{~~~~x++;} \\
		7 & \texttt{~~~~mutex\_unlock(\&m);} \\
		8 & \texttt{\}} \\
		%\texttt{~~~~\}} \\
		%\texttt{\}} \\
		\end{tabular}
	\end{center}
	\caption{Example transactional program.
		%The example {\tt count} routine from Figure~\ref{fig:mutex}, rewritten to use HTM.
		If the
		%transaction in the
		top branch aborts,
		%whether from a memory conflict or random system interrupt,
		%from the programmer's intention,
		execution will revert to the return of {\tt \_xbegin()}
		%{\tt status} will be assigned an error code
		and control will drop into the {\tt else} branch.
		The programmer can then use explicit synchronization
		%, such as a mutex,
		to resolve the conflict.}
	\label{fig:htm-example}
\end{figure}

@@

\definecolor{lavender}{RGB}{102,65,208} % (?? i don't know what happened here anymore but it looks good)
\definecolor{seafoam}{RGB}{65,178,76} % * 3/4
\definecolor{salmon}{RGB}{208,89,74} % * 7/8

@@

\begin{figure}[t]
	\begin{center}
		\begin{tabular}{ll}
		%\texttt{void count() \{} \\
		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
		  & \texttt{prevent\_transactions = false;} \\
		\\
		0 & \texttt{while (prevent\_transactions) continue;} \\
		1 & \texttt{if ((status = \_xbegin()) == SUCCESS) \{} \\
		2 & \texttt{~~~~if (prevent\_transactions)} \\
		3 & \texttt{~~~~~~~~\_xabort();} \\
		4 & \texttt{~~~~x++;} \\
		5 & \texttt{~~~~\_xend();} \\
		6 & \texttt{\} else \{} \\
		7 & \texttt{~~~~mutex\_lock(\&m);} \\
		8 & \texttt{~~~~prevent\_transactions = true;} \\
		9 & \texttt{~~~~x++;} \\
		A & \texttt{~~~~prevent\_transactions = false;} \\
		B & \texttt{~~~~mutex\_unlock(\&m);} \\
		C & \texttt{\}} \\
		%\texttt{~~~~\}} \\
		%\texttt{\}} \\
		\end{tabular}
	\end{center}
	\caption{Variant of the program in Figure~\ref{fig:htm-example},
		with additional synchronization to protect the failure path from the transactional path.
		The optional line 0 serves to prevent a cascade of failure paths
		for the sake of performance
		by allowing threads to wait until transacting is safe again.}
	\label{fig:htm-fixed}
\end{figure}

@@

\definecolor{darkcyan}{RGB}{0,128,128}
\definecolor{lime}{RGB}{48,128,0}
\definecolor{pinkish}{RGB}{128,34,102} % ff4488
\definecolor{brownish}{RGB}{128,96,0}
\definecolor{darklavender}{RGB}{68,44,138}
\newcommand\ETA[1]{\hilight{brownish}{{\em #1}}\xspace}
\newcommand\cpu[1]{\hilight{darkcyan}{{#1}}\xspace}
\newcommand\wtm[1]{\hilight{lime}{{#1}}\xspace}
\newcommand\ints[1]{\hilight{pinkish}{{#1}}\xspace}

@@

% TODO syntax hilighting
\begin{figure}[t]
	%\begin{center}
		\begin{tabular}{l}
		%\texttt{void count() \{} \\
		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
		\texttt{while (\_retry);} \\
		\texttt{if (\_xbegin() == SUCCESS) \{} \\
		\texttt{~~~~tie(\_root,inserted) = \_insert(\_root,n); } \\
		\texttt{~~~~\_xend();} \\
		\texttt{\} else \{} \\
		\texttt{~~~~pthread\_mutex\_lock(\&\_tree\_lock);} \\
		\texttt{~~~~\_retry = true;} \\
		\texttt{~~~~tie(\_root,inserted) = \_insert(\_root,n); } \\
		\texttt{~~~~\_retry = false;} \\
		\texttt{~~~~pthread\_mutex\_unlock(\&\_tree\_lock);} \\
		\texttt{\}} \\
		\end{tabular}
	%\end{center}
	%\caption{mario man had a very large house. or a castle.}
	\caption{Unmodified code from {\tt htmavl.cpp} with the previously-unknown bug found by Landslide.
		The transaction path fails to check {\tt \_retry},
		leading to data races and corruption just as in {\tt htm1}.
		}
	\label{fig:avlbug}
\end{figure}
	%\item Despite Quicksand's ability to optimize for finding bugs in overall smaller state spaces,
	%	this does not correlate with better CPU-time or even better wall-clock time.
	%	While possibly a consequence of the modest test size,
	%	this suggests future work should extend the Quicksand algorithm
	%	to prioritize by state space maximality as well as by state space estimate.
	%\item

>>>>

% threads
\definecolor{lavender}{RGB}{102,65,208} % (?? i don't know what happened here anymore but it looks good)
\definecolor{seafoam}{RGB}{65,178,76} % * 3/4
\definecolor{salmon}{RGB}{208,89,74} % * 7/8

% eval data
\definecolor{darkcyan}{RGB}{0,128,128}
\definecolor{lime}{RGB}{48,128,0}
\definecolor{darklavender}{RGB}{68,44,138}
\definecolor{brownish}{RGB}{128,96,0}
\definecolor{pinkish}{RGB}{128,34,102} % ff4488

% syntax hilighting
\definecolor{brown}{RGB}{170,85,0}
\newcommand\flow[1]{\hilight{brown}{#1}\xspace}
\newcommand\const[1]{\hilight{brickred}{#1}\xspace}
\newcommand\call[1]{\hilight{pinkish}{#1}}

\begin{figure}[t]
	\begin{center}
		\begin{tabular}{ll}
		%\texttt{void count() \{} \\
		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
		1 & \texttt{\flow{if} ((status = \call{\_xbegin}()) == \const{SUCCESS}) \{} \\
		2 & \texttt{~~~~x++;} \\
		3 & \texttt{~~~~\call{\_xend}();} \\
		4 & \texttt{\} \flow{else} \{} \\
		5 & \texttt{~~~~\call{mutex\_lock}(\&m);} \\
		6 & \texttt{~~~~x++;} \\
		7 & \texttt{~~~~\call{mutex\_unlock}(\&m);} \\
		8 & \texttt{\}} \\
		%\texttt{~~~~\}} \\
		%\texttt{\}} \\
		\end{tabular}
	\end{center}
	\caption{Example transactional program.
		%The example {\tt count} routine from Figure~\ref{fig:mutex}, rewritten to use HTM.
		If the
		%transaction in the
		top branch aborts,
		%whether from a memory conflict or random system interrupt,
		%from the programmer's intention,
		execution will revert to the return of {\tt \_xbegin()}
		%{\tt status} will be assigned an error code
		and control will drop into the {\tt else} branch.
		The programmer can then use explicit synchronization
		%, such as a mutex,
		to resolve the conflict.}
	\label{fig:htm-example}
\end{figure}

@@

@@

\begin{figure}[t]
	\begin{center}
		\begin{tabular}{ll}
		%\texttt{void count() \{} \\
		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
		  & \texttt{prevent\_transactions = \const{false};} \\
		\\
		0 & \texttt{\flow{while} (prevent\_transactions) \flow{continue};} \\
		1 & \texttt{\flow{if} ((status = \call{\_xbegin}()) == \const{SUCCESS}) \{} \\
		2 & \texttt{~~~~\flow{if} (prevent\_transactions)} \\
		3 & \texttt{~~~~~~~~\call{\_xabort}();} \\
		4 & \texttt{~~~~x++;} \\
		5 & \texttt{~~~~\call{\_xend}();} \\
		6 & \texttt{\} \flow{else} \{} \\
		7 & \texttt{~~~~\call{mutex\_lock}(\&m);} \\
		8 & \texttt{~~~~prevent\_transactions = \const{true};} \\
		9 & \texttt{~~~~x++;} \\
		A & \texttt{~~~~prevent\_transactions = \const{false};} \\
		B & \texttt{~~~~\call{mutex\_unlock}(\&m);} \\
		C & \texttt{\}} \\
		%\texttt{~~~~\}} \\
		%\texttt{\}} \\
		\end{tabular}
	\end{center}
	\caption{Variant of the program in Figure~\ref{fig:htm-example},
		with additional synchronization to protect the failure path from the transactional path.
		The optional line 0 serves to prevent a cascade of failure paths
		for the sake of performance
		by allowing threads to wait until transacting is safe again.}
	\label{fig:htm-fixed}
\end{figure}

@@

\newcommand\ETA[1]{\hilight{brownish}{{\em #1}}\xspace}
\newcommand\cpu[1]{\hilight{darkcyan}{{#1}}\xspace}
\newcommand\wtm[1]{\hilight{lime}{{#1}}\xspace}
\newcommand\ints[1]{\hilight{pinkish}{{#1}}\xspace}

@@

% TODO syntax hilighting
\begin{figure}[t]
	%\begin{center}
		\begin{tabular}{l}
		%\texttt{void count() \{} \\
		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
		\texttt{\flow{while} (\_retry);} \\
		\texttt{\flow{if} (\call{\_xbegin}() == \const{SUCCESS}) \{} \\
		\texttt{~~~~\hilight{brickred}{tie}(\_root,inserted) = \call{\_insert}(\_root,n); } \\
		\texttt{~~~~\call{\_xend}();} \\
		\texttt{\} \flow{else} \{} \\
		\texttt{~~~~\call{pthread\_mutex\_lock}(\&\_tree\_lock);} \\
		\texttt{~~~~\_retry = \const{true};} \\
		\texttt{~~~~\hilight{brickred}{tie}(\_root,inserted) = \call{\_insert}(\_root,n); } \\
		\texttt{~~~~\_retry = \const{false};} \\
		\texttt{~~~~\call{pthread\_mutex\_unlock}(\&\_tree\_lock);} \\
		\texttt{\}} \\
		\end{tabular}
	%\end{center}
	%\caption{mario man had a very large house. or a castle.}
	\caption{Unmodified code from {\tt htmavl.cpp} with the previously-unknown bug found by Landslide.
		The transaction path fails to check {\tt \_retry},
		leading to data races and corruption just as in {\tt htm1}.
		}
	\label{fig:avlbug}
\end{figure}
	%\item Despite Quicksand's ability to optimize for finding bugs in overall smaller state spaces,
	%	this does not correlate with better CPU-time or even better wall-clock time.
	%	While possibly a consequence of the modest test size,
	%	this suggests future work should extend the Quicksand algorithm
	%	to prioritize by state space maximality as well as by state space estimate.
	%\item
